---
title: "COVID19 Cases and Transportation Data in Montgomery County, MD"
author: "Emily Yamauchi"
date: "12/5/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(forecast)
library(xts)
```


Explore relationships between COVID19 daily cases and transportation data in Montgomery County, MD

## Public Transportation

```{r loadTransit}

transit = read.csv('data_clean/transit_data.csv')

```


#### Hypothesis 1: Bus ridership driven by COVID19 Cases  
Mobility decisions are impacted by both sentiment and mandates. Bus ridership would decrease as cases increase, and increase as cases decrease. 


```{r}

plt <- ggplot(transit, aes(x=ravg_cases_pct, y=ravg_bus_pct)) + geom_point() + geom_smooth(method="lm", se=FALSE)
plt + labs(y="Bus Ridership (Daily % Chg)", x="COVID19 Cases (Daily % Chg)", title="COVID19 Cases vs Bus Ridership (Daily % Change)", subtitle="Montgomery County, MD. Jun 2020 - Oct 2021")

```


```{r}

m_bus = lm(ravg_bus_pct ~ ravg_cases_pct, data=transit)
summary(m_bus)
aov1 = aov(ravg_bus_pct ~ ravg_cases_pct, data = transit)


```
Linear regression to test linear effect of COVID19 cases on bus ridership is not significant- $p-$value=`r unlist(summary(aov1))["Pr(>F)1"]`  

#### Hypothesis 2: Rail ridership driven by COVID19 Cases  
Mobility decisions are impacted by both sentiment and mandates. Rail ridership would decrease as cases increase, and increase as cases decrease. 

```{r}

plt <- ggplot(transit, aes(x=ravg_cases_pct, y=ravg_rail_pct)) + geom_point() + geom_smooth(method="lm", se=FALSE)
plt + labs(y="Rail Ridership (Daily % Chg)", x="COVID19 Cases (Daily % Chg)", title="COVID19 Cases vs Rail Ridership (Daily % Change)", subtitle="Montgomery County, MD. Jun 2020 - Oct 2021")

```


```{r }

m_rail = lm(ravg_rail_pct ~ ravg_cases_pct, data=transit)
summary(m_rail)
aov2 = aov(ravg_rail_pct ~ ravg_cases_pct, data = transit)

```

Linear regression to test linear effect of COVID19 cases on rail ridership is not significant- $p-$value=`r unlist(summary(aov2))["Pr(>F)1"]`  

#### Hypothesis 3: Transit ridership driven by COVID19 Cases  
Mobility decisions are impacted by both sentiment and mandates. Bus and rail ridership would decrease as cases increase, and increase as cases decrease. 

```{r}

plt <- ggplot(transit, aes(x=ravg_cases_pct, y=ravg_all_pct)) + geom_point() + geom_smooth(method="lm", se=FALSE)
plt + labs(y="Transit Ridership (Daily % Chg)", x="COVID19 Cases (Daily % Chg)", title="COVID19 Cases vs Transit Ridership (Daily % Change)", subtitle="Montgomery County, MD. Jun 2020 - Oct 2021")

```

```{r}

m_all = lm(ravg_all_pct ~ ravg_cases_pct, data=transit)
summary(m_all)
aov3 = aov(ravg_all_pct ~ ravg_cases_pct, data = transit)

```

Linear regression to test linear effect of COVID19 cases on total transit ridership is not significant- $p-$value=`r unlist(summary(aov3))["Pr(>F)1"]`  


## Collisions Data  

```{r loadCollisions}

crash = read.csv('data_clean/crashes_data.csv')

case_pct2 = crash[,"cases_pct"]
crash_pct = crash[,"crash_pct"]

```

#### Hypothesis 3: Transit ridership driven by COVID19 Cases  
Similar to Hypothesis 1, where mobility decisions are impacted by both sentiment and mandates. With less drivers on the road, number of collisions would decrease as cases increase.

```{r}

plt <- ggplot(crash, aes(x=case_pct2, crash_pct)) + geom_point() + geom_smooth(method="lm", se=FALSE)
plt + labs(y="Number of Collisions (Daily % Chg)", x="COVID19 Cases (Daily % Chg)", title="Collisions vs COVID19 Cases (Daily % Change)", subtitle="Montgomery County, MD. Jun 2020 - Aug 2021")

```



```{r}

m_crash = lm(crash_pct ~ case_pct2, data=crash)
summary(m_crash)
aov5 = aov(crash_pct ~ case_pct2, data = crash)


```

Linear regression to test linear effect of COVID19 cases on collisions is not significant- $p-$value=`r unlist(summary(aov5))["Pr(>F)1"]`  